\pkg{NeuralNetTools}: Visualization and Analysis Tools for Neural Networks

### 2015-09-30 nnt_manu.Rnw
better figures with subcaptions

\usepackage{acronym}
\usepackage{cleveref}
\acrodef{mlp}[MLP]{multilayer perceptron}
\acrodef{nid}[NID]{neural interpretation diagram}

@@

<<plotnet, fig.show = 'hold', fig.height = 5.5, fig.width = 5.5, out.height = '7.5cm', out.width = '7.5cm', fig.cap = 'Examples from the \\code{plotnet} function showing neural networks as a standard graphic (left) and using the \\acl{nid} (right).  Labels outside of the nodes represent variable names and labels within the nodes indicate the layer and node (I: input, H: hidden, O: output, B: bias).  Default options are changed for the example.'>>=
par(mar = c(0, 0, 0, 0))
plotnet(mod3, nid = F, circle_col = 'grey', bord_col = 'grey')
plotnet(mod3)
@

@@

<<plotnet_skip, fig.show = 'hold', fig.height = 5.5, fig.width = 5.5, out.height = '7.5cm', out.width = '7.5cm', fig.cap = 'Examples from the \\code{plotnet} function showing a neural network with a separate skip layer between the input and output layers.  The skip layer (left) and primary neural network (right) can be viewed separately with \\code{plotnet}.', results = 'hide'>>=
# create a model with a skip layer
modskip <- nnet(Y1 ~ X1 + X2 + X3, data = neuraldat, size = 5, skip = TRUE)

@@

<<plotprune, eval = T, echo = T, fig.show = 'hold', fig.height = 7, fig.width = 7.7, out.width = '0.48\\textwidth', fig.cap = 'A pruned neural network using the ``Optimal Brain Surgeon\'\' algorithm described in \\citet{Zell98}.  The default plotting behavior of \\code{plotnet} is to omit pruned connections (left), whereas they can be viewed as dashed lines by including the \\code{prune\\_col} argument (right).'>>=
# pruned model using RSNNS
pruneFuncParams <- list(max_pr_error_increase = 10.0, pr_accepted_error = 1.0,
  no_of_pr_retrain_cycles = 1000, min_error_to_stop = 0.01, 
  init_matrix_value = 1e-6, input_pruning = TRUE, hidden_pruning = TRUE)
mod <- mlp(x, y, size = 5, pruneFunc = "OptimalBrainSurgeon",
 pruneFuncParams = pruneFuncParams)

@@

Both functions have similar implementations and require only a model object as input.  The default output is a \pkg{ggplot2} bar plot \citep[i.e., \code{geom\_bar},][]{Wickham09} that shows the relative importance of each input variable in the model.  The plot aesthetics are based on internal code from the function but can be changed using conventional syntax for \pkg{ggplot2}.  The importance values can also be returned as a \code{data.frame} if \code{bar\_plot = FALSE}.

<<plotimp, fig.show = 'hold', fig.height = 2, fig.width = 4, out.width = '0.45\\textwidth', fig.cap = 'Variable importance'>>=
# garson and olden functions for each model
garson(mod1)
olden(mod1)
garson(mod2)
olden(mod2)
garson(mod3)
olden(mod3)
@

>>>>

\usepackage{acronym}
\usepackage{cleveref}
\usepackage{subfig}

\acrodef{mlp}[MLP]{multilayer perceptron}
\acrodef{nid}[NID]{neural interpretation diagram}

@@

<<plotnet, fig.show = 'hold', fig.height = 5.5, fig.width = 5.5, out.height = '7.5cm', out.width = '7.5cm', fig.cap = 'Examples from the \\code{plotnet} function showing neural networks as a standard graphic (\\ref{fig:plotnet1}) and using the \\acl{nid} (\\ref{fig:plotnet2}).  Labels outside of the nodes represent variable names and labels within the nodes indicate the layer and node (I: input, H: hidden, O: output, B: bias).  Default options are changed for the example.', fig.subcap = c('', '')>>=
par(mar = c(0, 0, 0, 0))
plotnet(mod3, nid = F, circle_col = 'grey', bord_col = 'grey')
plotnet(mod3)
@

@@

<<plotnet_skip, fig.show = 'hold', fig.height = 5.5, fig.width = 5.5, out.height = '7.5cm', out.width = '7.5cm', fig.cap = 'Examples from the \\code{plotnet} function showing a neural network with a separate skip layer between the input and output layers.  The skip layer (\\ref{fig:plotnet_skip1}) and primary neural network (\\ref{fig:plotnet_skip2}) can be viewed separately with \\code{plotnet}.', results = 'hide', fig.subcap = c('', '')>>=
# create a model with a skip layer
modskip <- nnet(Y1 ~ X1 + X2 + X3, data = neuraldat, size = 5, skip = TRUE)

@@

<<plotprune, eval = T, echo = T, fig.show = 'hold', fig.height = 7, fig.width = 7.7, out.width = '0.48\\textwidth', fig.cap = 'A pruned neural network using the ``Optimal Brain Surgeon\'\' algorithm described in \\citet{Zell98}.  The default plotting behavior of \\code{plotnet} is to omit pruned connections (\\ref{fig:plotprune1}), whereas they can be viewed as dashed lines by including the \\code{prune\\_col} argument (\\ref{fig:plotprune2}).', fig.subcap = c('', '')>>=
# pruned model using RSNNS
pruneFuncParams <- list(max_pr_error_increase = 10.0, pr_accepted_error = 1.0,
  no_of_pr_retrain_cycles = 1000, min_error_to_stop = 0.01, 
  init_matrix_value = 1e-6, input_pruning = TRUE, hidden_pruning = TRUE)
mod <- mlp(x, y, size = 5, pruneFunc = "OptimalBrainSurgeon",
 pruneFuncParams = pruneFuncParams)

@@

Both functions have similar implementations and require only a model object as input.  The default output is a \pkg{ggplot2} bar plot \citep[i.e., \code{geom\_bar},][]{Wickham09} that shows the relative importance of each input variable in the model (\cref{fig:plotimp}).  The plot aesthetics are based on internal code from the function but can be changed using conventional syntax for \pkg{ggplot2}.  The importance values can also be returned as a \code{data.frame} if \code{bar\_plot = FALSE}.

<<eval = F>>=
# garson and olden functions for each model
garson(mod1)
olden(mod1)
garson(mod2)
olden(mod2)
garson(mod3)
olden(mod3)
@

\begin{figure}
\centering
\subfloat[][Model 1 (\code{mlp}), \code{garson}]{
<<plotimp-1, fig.height = 2, fig.width = 4, out.width = '0.47\\textwidth', echo = F>>=
garson(mod1)
@
\label{fig:plotimp1}
}
\subfloat[][Model 1 (\code{mlp}), \code{olden}]{
<<plotimp-2, fig.height = 2, fig.width = 4, out.width = '0.47\\textwidth', echo = F>>=
olden(mod1)
@
\label{fig:plotimp2}
}

\subfloat[][Model 2 (\code{nn}), \code{garson}]{
<<plotimp-3, fig.height = 2, fig.width = 4, out.width = '0.47\\textwidth', echo = F>>=
garson(mod2)
@
\label{fig:plotimp3}
}
\subfloat[][Model 2 (\code{nn}), \code{olden}]{
<<plotimp-4, fig.height = 2, fig.width = 4, out.width = '0.47\\textwidth', echo = F>>=
olden(mod2)
@
\label{fig:plotimp4}
}

\subfloat[][Model 3 (\code{nnet}), \code{garson}]{
<<plotimp-5, fig.height = 2, fig.width = 4, out.width = '0.47\\textwidth', echo = F>>=
garson(mod3)
@
\label{fig:plotimp5}
}
\subfloat[][Model 3 (\code{nnet}), \code{olden}]{
<<plotimp-6, fig.height = 2, fig.width =4, out.width = '0.47\\textwidth', echo = F>>=
olden(mod3)
@
\label{fig:plotimp6}
}
\caption{Variable importance for three models using Garson's algorithm for relative importance (\code{garson}, \citealt{Garson91,Goh95}) and Olden's conection weights algorithm (\code{olden}, \citealt{Olden04}).  Garson's algorithm shows importance as absolute values from 0--1 whereas Olden's algorithm preserves sign and magnitude.  Importance values for Olden's algorithm are directly from the summed product of model weights and should only be evaluated based on the sign and relative magnitudes between variables.}
\label{fig:plotimp}
\end{figure}

### 2015-09-30 nnt_manu.tex
better figures with subcaptions

{\centering \includegraphics[width=7.5cm,height=7.5cm]{figs/plotnet-1} 
\includegraphics[width=7.5cm,height=7.5cm]{figs/plotnet-2}

@@

\caption[Examples from the \code{plotnet} function showing neural networks as a standard graphic (left) and using the \acl{nid} (right)]{Examples from the \code{plotnet} function showing neural networks as a standard graphic (left) and using the \acl{nid} (right).  Labels outside of the nodes represent variable names and labels within the nodes indicate the layer and node (I: input, H: hidden, O: output, B: bias).  Default options are changed for the example.}\label{fig:plotnet}
\end{figure}

@@

{\centering \includegraphics[width=7.5cm,height=7.5cm]{figs/plotnet_skip-1} 
\includegraphics[width=7.5cm,height=7.5cm]{figs/plotnet_skip-2}

@@

\caption{Examples from the \code{plotnet} function showing a neural network with a separate skip layer between the input and output layers.  The skip layer (left) and primary neural network (right) can be viewed separately with \code{plotnet}.}\label{fig:plotnet_skip}
\end{figure}

@@

{\centering \includegraphics[width=0.48\textwidth]{figs/plotprune-1} 
\includegraphics[width=0.48\textwidth]{figs/plotprune-2}

@@

\caption{A pruned neural network using the ``Optimal Brain Surgeon'' algorithm described in \citet{Zell98}.  The default plotting behavior of \code{plotnet} is to omit pruned connections (left), whereas they can be viewed as dashed lines by including the \code{prune\_col} argument (right).}\label{fig:plotprune}
\end{figure}

@@

Both functions have similar implementations and require only a model object as input.  The default output is a \pkg{ggplot2} bar plot \citep[i.e., \code{geom\_bar},][]{Wickham09} that shows the relative importance of each input variable in the model.  The plot aesthetics are based on internal code from the function but can be changed using conventional syntax for \pkg{ggplot2}.  The importance values can also be returned as a \code{data.frame} if \code{bar\_plot = FALSE}.

\begin{kframe}
\begin{alltt}
\hlcom{# garson and olden functions for each model}
\hlkwd{garson}\hlstd{(mod1)}
\hlkwd{olden}\hlstd{(mod1)}
\hlkwd{garson}\hlstd{(mod2)}
\hlkwd{olden}\hlstd{(mod2)}
\hlkwd{garson}\hlstd{(mod3)}
\hlkwd{olden}\hlstd{(mod3)}
\end{alltt}
\end{kframe}\begin{figure}[!ht]

{\centering \includegraphics[width=0.45\textwidth]{figs/plotimp-1} 
\includegraphics[width=0.45\textwidth]{figs/plotimp-2} 
\includegraphics[width=0.45\textwidth]{figs/plotimp-3} 
\includegraphics[width=0.45\textwidth]{figs/plotimp-4} 
\includegraphics[width=0.45\textwidth]{figs/plotimp-5} 
\includegraphics[width=0.45\textwidth]{figs/plotimp-6}

@@

\caption[Variable importance]{Variable importance}\label{fig:plotimp}
\end{figure}

>>>>

{\centering \subfloat[\label{fig:plotnet1}]{\includegraphics[width=7.5cm,height=7.5cm]{figs/plotnet-1} }
\subfloat[\label{fig:plotnet2}]{\includegraphics[width=7.5cm,height=7.5cm]{figs/plotnet-2} }

@@

\caption{Examples from the \code{plotnet} function showing neural networks as a standard graphic (\ref{fig:plotnet1}) and using the \acl{nid} (\ref{fig:plotnet2}).  Labels outside of the nodes represent variable names and labels within the nodes indicate the layer and node (I: input, H: hidden, O: output, B: bias).  Default options are changed for the example.}\label{fig:plotnet}
\end{figure}

@@

{\centering \subfloat[\label{fig:plotnet_skip1}]{\includegraphics[width=7.5cm,height=7.5cm]{figs/plotnet_skip-1} }
\subfloat[\label{fig:plotnet_skip2}]{\includegraphics[width=7.5cm,height=7.5cm]{figs/plotnet_skip-2} }

@@

\caption{Examples from the \code{plotnet} function showing a neural network with a separate skip layer between the input and output layers.  The skip layer (\ref{fig:plotnet_skip1}) and primary neural network (\ref{fig:plotnet_skip2}) can be viewed separately with \code{plotnet}.}\label{fig:plotnet_skip}
\end{figure}

@@

{\centering \subfloat[\label{fig:plotprune1}]{\includegraphics[width=0.48\textwidth]{figs/plotprune-1} }
\subfloat[\label{fig:plotprune2}]{\includegraphics[width=0.48\textwidth]{figs/plotprune-2} }

@@

\caption{A pruned neural network using the ``Optimal Brain Surgeon'' algorithm described in \citet{Zell98}.  The default plotting behavior of \code{plotnet} is to omit pruned connections (\ref{fig:plotprune1}), whereas they can be viewed as dashed lines by including the \code{prune\_col} argument (\ref{fig:plotprune2}).}\label{fig:plotprune}
\end{figure}

@@

Both functions have similar implementations and require only a model object as input.  The default output is a \pkg{ggplot2} bar plot \citep[i.e., \code{geom\_bar},][]{Wickham09} that shows the relative importance of each input variable in the model (\cref{fig:plotimp}).  The plot aesthetics are based on internal code from the function but can be changed using conventional syntax for \pkg{ggplot2}.  The importance values can also be returned as a \code{data.frame} if \code{bar\_plot = FALSE}.

\begin{kframe}
\begin{alltt}
\hlcom{# garson and olden functions for each model}
\hlkwd{garson}\hlstd{(mod1)}
\hlkwd{olden}\hlstd{(mod1)}
\hlkwd{garson}\hlstd{(mod2)}
\hlkwd{olden}\hlstd{(mod2)}
\hlkwd{garson}\hlstd{(mod3)}
\hlkwd{olden}\hlstd{(mod3)}
\end{alltt}
\end{kframe}

\begin{figure}
\centering
\subfloat[][Model 1 (\code{mlp}), \code{garson}]{

{\centering \includegraphics[width=0.47\textwidth]{figs/plotimp-1-1}

@@

\label{fig:plotimp1}
}
\subfloat[][Model 1 (\code{mlp}), \code{olden}]{

{\centering \includegraphics[width=0.47\textwidth]{figs/plotimp-2-1}

\label{fig:plotimp2}
}

\subfloat[][Model 2 (\code{nn}), \code{garson}]{

{\centering \includegraphics[width=0.47\textwidth]{figs/plotimp-3-1}

\label{fig:plotimp3}
}
\subfloat[][Model 2 (\code{nn}), \code{olden}]{

{\centering \includegraphics[width=0.47\textwidth]{figs/plotimp-4-1}

\label{fig:plotimp4}
}

\subfloat[][Model 3 (\code{nnet}), \code{garson}]{

{\centering \includegraphics[width=0.47\textwidth]{figs/plotimp-5-1}

\label{fig:plotimp5}
}
\subfloat[][Model 3 (\code{nnet}), \code{olden}]{

{\centering \includegraphics[width=0.47\textwidth]{figs/plotimp-6-1}

\label{fig:plotimp6}
}
\caption{Variable importance for three models using Garson's algorithm for relative importance (\code{garson}, \citealt{Garson91,Goh95}) and Olden's conection weights algorithm (\code{olden}, \citealt{Olden04}).  Garson's algorithm shows importance as absolute values from 0--1 whereas Olden's algorithm preserves sign and magnitude.  Importance values for Olden's algorithm are directly from the summed product of model weights and should only be evaluated based on the sign and relative magnitudes between variables.}
\label{fig:plotimp}
\end{figure}
